---
title: Seeing is Conceiving
author: "Brian A. Fannin"
date: August 3, 2019
output: 
  revealjs::revealjs_presentation:
    center: false
    transition: none
    css: css/custom.css
    self_contained: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE
  , warning = FALSE
  , message = FALSE
  , collapse = TRUE
  , cache = TRUE
)

library(datasauRus)
library(gridExtra)
library(tidyverse)
```

```{r}
my_void <- theme_void() +
  theme(
    text = element_text(size = 20)
    , legend.position = 'none'
  )

my_minimal <- theme_minimal() +
  theme(
    text = element_text(size = 20)
  )

my_classic <- theme_classic() + 
  theme(
    text = element_text(size = 20)
  )

my_percent <- function(x) {
  scales::percent(x, accuracy = 0.1)
}

my_args <- list(big.mark = ',', digits = 0, scientific = FALSE)
```

```{r}
load('data/data.rda')
```

# Overview

## {.center}

Numbers are more than symbols

Visualization == information density

# Numbers are more than symbols

## {.center .very-large-text-full}

Seeing is believing
<!--
## Scaling the Zen mountain 

<div class = 'center'>
“Before I studied Zen, I saw mountains as mountains and rivers as rivers. When I had studied Zen for thirty years I no longer saw mountains as mountains and rivers as rivers. But now that I have finally mastered Zen, I once again see mountains as mountains and rivers as rivers.”

> Ch’an master Ch’ing Yuan

</div>
-->

## {.center}

conceive (v) - form a mental representation of; imagine.

## {.center}

Seeing is _conceiving_

## {.center}

conceive (v) - form a mental representation of; imagine.

## {data-background=white}

## {.gigantic-text}

9 

## {.center}

Which of these two numbers is larger? 

##

<div class = 'left very-large-text'>
11
</div>

<div class = 'right very-large-text'>
9
</div>

## {.center}

How about these two?

##

<div class = 'left large-text'>
1011
</div>

<div class = 'right large-text'>
1001
</div>

## {.center}

And these?

##

<div class = 'left very-large-text'>
IX
</div>

<div class = 'right very-large-text'>
XI
</div>

## {.center}

These?

##

<div class = 'left large-text'>
十一
</div>

<div class = 'right large-text'>
九
</div>

## {.center}

How about these?

##

```{r}
tibble(
    val = c('left', 'right')
  , size = c(11, 9)
) %>% 
  ggplot(
    aes(val, size)
  ) + 
  geom_bar(position = 'dodge', stat = 'identity') + 
  my_void
```

## {.center}

nine

9

neun

1001

九

IX

nueve

##  {.center}

Arabic or sanskrit are no more legitimate than any other representation of numbers.

## {.center}

9 != nine

9 <> nine

## {.center}

_We can no more see numbers than we can taste or smell them._

# Visualization == information density

## This is actually _too much_ information

```{r}
tibble(
    val = c('left', 'right')
  , size = c(11, 9)
) %>% 
  ggplot(
    aes(val, size)
  ) + 
  geom_bar(position = 'dodge', stat = 'identity') + 
  my_void
```

## This is better

```{r fig.width=4}
tibble(
    val = c('left', 'right')
  , size = c(11, 9)
) %>% 
  ggplot(
    aes(val, size)
  ) + 
  geom_bar(position = 'dodge', stat = 'identity', width = 0.1) + 
  my_void
```

##

Statistics maps a set of many numbers into a set of fewer numbers.

```{r}
set.seed(1234)
mean <- 10e3
cv <- 0.3
sigma <- sqrt(log(1 + cv^2))
mu <- log(mean) - sigma^2/2

tbl_obs <- tibble(
  x = rlnorm(5e3, mu, sigma)
  , y = rgamma(5e3, 1 / cv^2, scale = mean * cv ^ 2)
)

tbl_obs_long <- tbl_obs %>% 
  tidyr::gather(sample, value)

tbl_obs <- tbl_obs %>% 
  mutate(
    yen = x * 137
  )

summary_x <- tbl_obs$x %>% 
  summary()

summary_y <- tbl_obs$y %>% 
  summary()

summary_yen <- tbl_obs$yen %>% 
  summary()

tbl_summary <- data.frame(
  metric = names(summary_x)
  , x = summary_x %>% as.numeric()
  , y = summary_y %>% as.numeric()
  , yen = summary_yen %>% as.numeric()
)
```

```{r}
mean_sample <- tbl_obs$x %>% mean()
sd_sample <- tbl_obs$x %>% sd()
```

```{r }
tbl_obs$x %>% 
  summary()
```

## {.center}

```{r }
tbl_summary %>%
  select(metric, x) %>% 
  knitr::kable(
    format.args = my_args
  )
```

## 

```{r}
plt_base <- tbl_obs %>% 
  ggplot(aes(x))

plt_hist <- plt_base + 
  geom_histogram(
      aes(y = stat(density))
    , fill = 'grey'
    , color = 'black')

plt_density <- plt_base + 
  geom_density(fill = 'grey')

grid.arrange(
    nrow = 1
  , plt_hist
  , plt_density
)
```

## 

```{r}
mean_x <- tbl_obs$x %>% mean()

grid.arrange(
    nrow = 1
  , plt_hist + geom_vline(xintercept = mean_x, color = 'red')
  , plt_density + geom_vline(xintercept = mean_x, color = 'red')
)
```

##

```{r}
conf_95 <- quantile(tbl_obs$x, c(.05, .95))

grid.arrange(
    nrow = 1
  , plt_hist + 
      geom_vline(xintercept = mean_x, color = 'red') + 
      geom_vline(xintercept = conf_95, color = 'black', alpha = 0.6)
  , plt_density + geom_vline(xintercept = mean_x, color = 'red')  + 
      geom_vline(xintercept = conf_95, color = 'black', alpha = 0.6)
)
```

##

<div class = 'left'>

```{r }
tbl_summary %>% 
  select(metric, x) %>% 
  knitr::kable(
    format.args = my_args
  )
```

</div>

<div class = 'right'>

```{r fig.height = 8}
plt_hist + 
  geom_vline(xintercept = mean_x, color = 'red') + 
  geom_vline(xintercept = conf_95, color = 'black', alpha = 0.6)
```

</div>

## {.center}

Summary statistics are _always_ a reduction of information.

Visualization presents (almost) _all_ of the data. The reductions are made with our eyes. 

## {.center}

Scale transforms

## {.center}

```{r}
tbl_summary %>% 
  select(metric, x, z = yen) %>% 
  knitr::kable(
    format.args = my_args
  )
```

##

```{r}
plt_x <- tbl_obs %>% 
  ggplot(aes(x)) + 
  geom_histogram(
      aes(y = stat(density))
    , fill = 'grey'
    , color = 'black')

plt_yen <- tbl_obs %>% 
  ggplot(aes(yen)) + 
  geom_histogram(
      aes(y = stat(density))
    , fill = 'grey'
    , color = 'black')

grid.arrange(
    nrow = 1
  , plt_x
  , plt_yen
)
```

## {.center}

Visualization as a metaphor for a model

## {.center}

```{r}
tbl_summary %>% 
  select(metric, x, y) %>% 
  knitr::kable(
    format.args = my_args
  )
```

## 

```{r}
plt_ln_base <- tbl_obs_long %>% 
  filter(sample == 'x') %>% 
  ggplot(aes(sample, value)) +
  geom_boxplot()

grid.arrange(
  nrow = 1
  , plt_ln_base
  , plt_ln_base + scale_y_log10()
)
```

##

```{r}
tbl_obs_long %>% 
  ggplot(aes(sample, value)) + 
  geom_boxplot() + 
  scale_y_log10()
```

## Financial reporting view

```{r}
tbl_policy_uy <- tbl_policy %>% 
  group_by(uy = lubridate::year(effective_date)) %>% 
  summarise(
    mean_claims = mean(num_claims)
  )
```

```{r }
tbl_policy_uy %>% 
  knitr::kable(format.args = list(digits = 2))
```

##

```{r}
tbl_policy_uy %>% 
  ggplot(aes(uy, mean_claims)) + 
  geom_point() + 
  geom_smooth(method = 'lm') + 
  scale_x_continuous(breaks = tbl_policy_uy$uy %>% unique())
```

##

```{r}
tbl_policy %>% 
  ggplot(aes(effective_date, num_claims)) + 
  geom_point() + 
  geom_smooth(method = 'lm')
```

##

```{r}
tbl_policy %>%
  mutate(
    uy = lubridate::year(effective_date)
  ) %>% 
  ggplot(aes(uy, num_claims)) + 
  geom_boxplot(aes(group = uy)) + 
  geom_smooth(method = 'lm') + 
  scale_x_continuous(breaks = tbl_policy_uy$uy %>% unique())
```

##

<div class = 'left'>

```{r}
tbl_policy_uy %>% 
  knitr::kable(format.args = list(digits = 2))
```

</div>

<div class = 'left'>

```{r fig.height=8}
tbl_policy %>%
  mutate(
    uy = lubridate::year(effective_date)
  ) %>% 
  ggplot(aes(uy, num_claims)) + 
  geom_boxplot(aes(group = uy)) + 
  geom_smooth(method = 'lm') + 
  scale_x_continuous(breaks = tbl_policy_uy$uy %>% unique())

```

</div>

<!--
##

```{r}
tbl_policy %>% 
  ggplot(aes(num_claims)) + 
  geom_histogram(aes(fill = region), position = 'dodge', binwidth = 1) + 
  facet_wrap(~ region)
```

## 

```{r}
tbl_policy %>% 
  ggplot(aes(num_claims)) + 
  geom_density(aes(fill = region), alpha = 0.7)
```
-->

## Validating models

A good model is one which isolates signal and leaves only noise.  

## 

```{r}
data(anscombe)

tbl_anscombe <- anscombe %>% 
  tidyr::gather() %>% 
  tidyr::separate(key, c('var', 'segment'), sep = 1) %>% 
  group_by(var, segment) %>% 
  mutate(index = seq_len(n())) %>% 
  ungroup() %>% 
  tidyr::spread(var, value) %>% 
  mutate(
    segment = as.integer(segment)
  )
```

```{r}
fit_segment <- function(which_segment) {
  tbl_anscombe %>% 
    dplyr::filter(segment == which_segment) %>% 
    lm(formula = y ~ 1 + x)
}

tbl_fit <- tbl_anscombe %>% 
  group_by(segment) %>% 
  summarise(
    n = n()
  ) %>% 
  mutate(
    fit = map(segment, fit_segment)
  )

tbl_model <- map_dfr(tbl_fit$fit, broom::glance) %>% 
  mutate(segment = 1:4)
```

```{r}
tbl_model %>% 
  dplyr::select(segment, adj.r.squared, sigma) %>% 
  mutate(
    adj.r.squared = my_percent(adj.r.squared)
    , sigma = format(sigma, digits = 2)
  ) %>% 
  knitr::kable()
```

Are these good models?

## 

```{r }
tbl_anscombe %>% 
  ggplot(aes(x, y)) + 
  geom_point() + 
  facet_wrap(~segment, scales = 'free') + 
  geom_smooth(method = 'lm', se = FALSE) + 
  my_minimal
```

## 

```{r}
tbl_residuals <- map_dfr(tbl_fit$fit, broom::augment) %>% 
  mutate(
    segment = rep(1:4, 11) %>% sort()
  ) %>% 
  rename(
    fitted = `.fitted`
    , residual = `.std.resid`
  )


tbl_residuals %>% 
  ggplot(aes(fitted, residual)) + 
  geom_point() + 
  facet_wrap(~ segment, scales = 'free') + 
  geom_hline(yintercept = 0, color = 'blue') + 
  geom_hline(yintercept = c(3, -3), color = 'red', linetype = 'dashed') + 
  my_minimal
```

## {.center .very-large-text-full}

Correlation is not causation

##

```{r}
data(datasaurus_dozen)

tbl_cor <- datasaurus_dozen %>% 
  group_by(dataset) %>% 
  summarise(
    correlation = cor(x, y)
  ) %>% 
  mutate(set = letters[seq_len(nrow(.))])
```

```{r}
tbl_cor %>% 
  dplyr::select(set, correlation) %>% 
  mutate(correlation = scales::percent(correlation, accuracy = .1)) %>% 
  knitr::kable()
```

##

```{r}
this_cor <- tbl_cor$correlation[tbl_cor$dataset == 'away'] %>% 
  scales::percent(accuracy = 0.1)

datasaurus_dozen %>% 
  filter(dataset == 'away') %>% 
  ggplot(aes(x, y)) + 
  geom_point() + 
  labs(title = paste('Correlation = ', this_cor)) + 
  my_classic
```

## {.center}

Another

##

```{r}
this_cor <- tbl_cor$correlation[tbl_cor$dataset == 'slant_up'] %>% 
  scales::percent(accuracy = 0.1)

datasaurus_dozen %>% 
  filter(dataset == 'slant_up') %>% 
  ggplot(aes(x, y)) + 
  geom_point() + 
  labs(title = paste('Correlation = ', this_cor)) + 
  my_classic
```

## {.center}

Another

## 

```{r}
this_cor <- tbl_cor$correlation[tbl_cor$dataset == 'dino'] %>% 
  scales::percent(accuracy = 0.1)

datasaurus_dozen %>% 
  filter(dataset == 'dino') %>% 
  ggplot(aes(x, y)) + 
  geom_point() + 
  labs(title = paste('Correlation = ', this_cor)) + 
  my_classic
```

# Conclusion

## {.center}

Numbers are more than symbols

Visualization == information density

## {.center .very-large-text-full}

Seeing is conceving

## Thank you!

## Where to find this

This presentation may be found at: http://pirategrunt.com/iaba_visualization

Code to produce the examples and slides: https://github.com/PirateGrunt/iaba_visualization 